Uncertainty propagation for predictor effects and bias correction in generalized linear (mixed) models

### 2023-02-08 paper/paper.tex
Ready for evening session

\begin{figure}
\begin{center}
\includegraphics{eco.splash.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines). The panel on the left anchors the effect-style intervals at the model center; the panel on the right has a specific anchor chosen for biological reasons. Anchor point does not affect prediction-style confidence intervals}
\flab{splash}
\end{figure}

@@

\subsection{The model center}

To make an outcome plot, we vary the value of the focal predictor, but we need to choose values for other predictors. Our default is to do this using the \emph{model center}. We define the model center as the average value of all of the “linear predictor variables” which go into the model matrix. This is not, in general, the same as averaging over the “input variables” which are measured and provided as data (see Glossary box).

@@

\begin{figure}
\begin{center}
\includegraphics{eco.multiPK.defAnchor.centers.plot.Rout.pdf}
\end{center}
\caption{Outcome plots for a model with a non-focal interaction. Black lines show an effect plot (mean estimate and CIs) made using the model center (average of linear predictor variables) as a reference point. Blue lines show the same approach made by centering based on input variables; due to the interaction this model does not go through the model center.}
\flab{centers}
\end{figure}

@@

\subsection{Direct effects}

\begin{figure}
\begin{center}
\includegraphics{eco.uniMulti.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines). The panel on the left anchors the effect-style intervals at the model center; the panel on the right has a specific anchor chosen for biological reasons. Anchor point does not affect prediction-style confidence intervals}
\flab{uniMulti}
\end{figure}

\fref{uniMulti} shows an example, based on simulated data with an interaction between correlated variables.

\section{Generalized linear models}

>>>>

\begin{figure}
\begin{center}
\includegraphics{eco.splash.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines) \CIs. The panel on the left anchors the effect-style intervals at the model center; the panel on the right has a specific anchor chosen for biological reasons. Anchor point does not affect prediction-style confidence intervals. Points show the fitted data; the large grey point shows the mean values of predictor and the response.}
\flab{splash}
\end{figure}

@@

\subsection{Direct and indirect effects}

\begin{figure}
\begin{center}
\includegraphics{eco.uniMulti.Rout.pdf}
%%
\caption{Outcome plots with effect- (dotted straight lines) and prediction-style (dashed curved lines). The panel on the left shows the results of a univariate model fit; the panel on the right shows a multivariate fit (and effect-style lines only, see text.}
%%
\flab{uniMulti}
\end{center}
\end{figure}
%%
\jd{Still working on this. One panel or two?}

\fref{uniMulti} shows an example.

\subsection{Reference points}

To make an outcome plot, we vary the value of the focal predictor, but we need to choose values for other predictors -- the reference point for the plot. Our default is to do this using the \emph{model center}. We define the model center as the average value of all of the “linear predictor variables” which go into the model matrix. This is not, in general, the same as averaging over the “input variables” which are measured and provided as data (see Glossary box).

@@

\begin{figure}
\begin{center}
\includegraphics{eco.multiPK.defAnchor.centers.plot.Rout.pdf}
\end{center}
\caption{Outcome plots for a model with a non-focal interaction. Black lines show an effect plot (mean estimate and CIs) made using the model center (average of linear predictor variables) as a reference point. Blue lines show the same approach made by centering based on input variables; due to the interaction this model does not go through the center point.}
\flab{centers}
\end{figure}

@@

An example is shown in \fref{quadComp}. We fit a univariate, \emph{quadratic} response to nitrogen for the biomass. Both panels show the same model, and have the same response curves and prediction CIs. The panel on the left shows the default effect CIs, anchored at the model center. This does not correspond to any real value of nitrogen, because the mean of the quadratic values is taken separately, and is not the value that would correspond in that column to the mean of the original values. FOOTNOTE about Jensen, or what? The panel on the right calculates effect CIs starting from the point that would correspond to an actual nitrogen value (in this case, the mean across the sampled values). The x-axis also reflects actual possible values of nitrogen; this is why the effect uncertainty goes to zero (the anchor point is visible) in the right panel but not the left panel. This is also why the response curves do not pass through the model-center point. 
\section{Generalized linear models}
